---
title: "linear models"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
library(knitr)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)
```

Bootstrapping helps with hypothesis testing when you're outside of a normal distribution. 

Regression diagnostics
- getting your model, looking at residuals, and seeing if it looks okay. 

Generalized linear models - logistic regression 

Broom package to tidy up linear model stuff 

## Start to code 

Load NYC Airbnb data

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location/2) %>% 
  rename(borough = neighbourhood_group) %>% 
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighbourhood, room_type)
```

Visualizations

```{r}
nyc_airbnb %>% 
  ggplot(aes(x = stars, y = price)) + 
  geom_point()
```

We have some assumptions that aren't met (there's more variability in price among higher stars than among lower stars)

Let's fit a linear model 

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

Let's look at this

```{r}
fit

summary(fit)

summary(fit)$coef

fit %>% broom::tidy()
```

Turns summary of fit into a dataframe 

If you want to present output... 

```{r}
fit %>% broom::tidy() %>% 
  mutate(term = str_replace(term, "borough", "Borough: ")) %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)
```

## Diagnostics 

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = stars, y = resid)) + 
  geom_point()
```

Returns a daataframe but has added in the residuals column
- could use this to compare residuals across many models 

Looks very skewed 

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = resid)) + 
  geom_density() + 
  xlim(-200, 200)
```

Because the whole distribution goes out to 10,000 it looks like you're missing a bunch of smaller values 

## Interactions? Nesting?

Let's try a different model ...

```{r}
fit = lm(price ~ stars + room_type, data = nyc_airbnb)

broom::tidy(fit)

fit = lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)

broom::tidy(fit)
```

Let's try nesting 

```{r}
nyc_airbnb %>% 
  relocate(borough) %>% 
  nest(data = price:room_type) %>% 
  mutate(
    lm_fits = map(.x = data, ~lm(price ~ stars + room_type, data = .x)), 
    lm_results = map(lm_fits, broom::tidy)) %>% 
  select(borough, lm_results) %>% 
  unnest(lm_results) %>% 
  filter(term == "stars")

```

One more of these examples - Neighborhoods in Manhattan 

```{r}
manhattan_lm_results_df = 
  nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  select(-borough) %>% 
  relocate(neighbourhood) %>% 
  nest(data = price:room_type) %>% 
  mutate(
    lm_fits = map(.x = data, ~lm(price ~ stars + room_type, data = .x)), 
    lm_results = map(lm_fits, broom::tidy)
  ) %>% 
  select(neighbourhood, lm_results) %>% 
  unnest(lm_results)

manhattan_lm_results_df %>% 
  filter(term == "stars") %>% 
  ggplot(aes(x = estimate)) + 
  geom_density()

manhattan_lm_results_df %>% 
  filter(str_detect(term, "room_type")) %>% 
  ggplot(aes(x = neighbourhood, y = estimate)) + 
  geom_point() +
  facet_grid(~term) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

It seems like stars don't really help that much (maybe more of the effect is room type or something else)

NEXT ONE: Keeping star rating fixed, going from Entire home to a Private room will be -100 (same with shared room - -120/150 maybe)

## Logistic regression 

```{r}
new_dataset = 
  nyc_airbnb %>% 
  mutate(
    expensive_apt = as.numeric(price > 500)
  )
```

Logistic regression 

```{r}
logistic_fit = 
  glm(
      expensive_apt ~ stars + borough, 
      data = new_dataset, 
      family = binomial())

logistic_fit %>% 
  broom::tidy() %>% 
  mutate(
    term = str_replace(term, "borough", "Borough: "), 
    estimate = exp(estimate)
  ) %>% 
  select(term, OR = estimate, p.value)

new_dataset %>% 
  modelr::add_predictions(logistic_fit) %>% 
  mutate(pred = boot::inv.logit(pred))
```






